Velocity Map Imaging Spectroscopy of C$_2$H$^-$ and C$_2$D$^-$: a benchmark study of vibronic coupling interactions

### 2022-04-07 Aug2021-CH2-.tex
sanov update

\documentclass[journal=jpcafh,manuscript=article,layout=onecolumn, 12pt]{achemso}
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage[T1]{fontenc}       % Use modern font encodings
\usepackage{amsmath}
\usepackage[dvipsnames]{xcolor}
% NB added command for in line cite
\newcommand{\onlinecite}[1]{\hspace{-1 ex} \nocite{#1}\citenum{#1}} 
%
\author{Benjamin~A.~Laws}
\email{b.laws@unsw.edu.au}
\affiliation{School of Chemistry, University of New South Wales, Sydney NSW 2052, Australia}
\alsoaffiliation{Research School of Physics, The Australian
	National University, Canberra ACT 2601, Australia}
\author{Zachariah~D.~Levey} 
\affiliation{School of Chemistry, University of New South Wales, Sydney NSW 2052, Australia}
\author{John F. Stanton}
\affiliation{Department of Chemistry, University of Florida, Gainesville, Florida 32611, United States}
\author{Timothy~W.~Schmidt} 
\affiliation{School of Chemistry, University of New South Wales, Sydney NSW 2052, Australia}
\author{Stephen~T.~Gibson}
\affiliation{Research School of Physics, The Australian
	National University, Canberra ACT 2601, Australia}
\title{Velocity Map Imaging Spectroscopy of C$_2$H$^-$ and C$_2$D$^-$: a benchmark study of vibronic coupling interactions}
\abbreviations{PES,PAD,EA,eKE,FWHM,VMI}
\begin{document} 
\begin{abstract}

@@

In this paper we study the vibronic spectrum of the ethynyl radical C$_2$H by employing High-Resolution Photoelectron Imaging (HR-PEI) in order to benchmark CFOUR~\cite{dev20}  \emph{ab-initio} vibronic-coupling calculations. C$_2$H is reported to be one of the most abundant molecules in the universe, and is the most thoroughly studied of the C$_{2n}$H species~\cite{wil91,hei99,for10}. Many different experimental techniques have been employed to try and resolve the complex vibronic spectrum, including electron spin resonance~\cite{coc64,jin85}, laser magnetic resonance~\cite{bro88,pfe96,sch98}, microwave and milli-meter wave spectroscopy~\cite{sas81,got83,end89,mul00}, infrared (matrix isolation~\cite{she87,jac87,for95} and Fourier Transform emission~\cite{ver88}) spectroscopy, laser induced fluorescence spectroscopy~\cite{hsu93,hsu95,chi99} and photoelectron spectroscopy~\cite{erv91,tay98,zho07}. C$_2$H has also received extensive theoretical attention to try and understand the experimental results, however the large number of possible vibronic levels means this has proven to be a challenge~\cite{per90,per91,per91b,per91c,per92,car00,tar03,tar04,for10,sta21}. In this work we demonstrate how construction of a quasidiabatic Hamiltonian allows for the strength of vibronic interactions between coupled surfaces near a conical intersection to be calculated, in order to simulate electronic and vibronic spectra. By employing anion HR-PEI both the $^2\Sigma^+$ and $^2\Pi$ surfaces are mapped out on an equal footing from the anion $\tilde{X}^1\Sigma^+$ state, allowing for direct comparison to the \emph{ab-initio} modelling.

\section{Results and Analysis}
Ethynyl ions were produced in a pulsed-jet discharge of pure C$_2$H$_4$ gas, and subsequently mass isolated via time of flight. Electrons were detached using a tuneable Sunlite Optical Parametric Oscillator (OPO) pumped with the third harmonic of a Nd:YAG laser. High electron counts could also be obtained by using the third and fourth harmonics of the Nd:YAG laser directly. The detached electrons were then mapped onto a micro channel plate detector using a Velocity Map Imaging (VMI) lens. An illustrative VMI of $\sim$4 million electrons collected from 355~nm (3.49~eV) photodetachment of C$_2$H$^-$ is shown in Fig.~\ref{fig:1}(a). The electrons are distributed radially according to their speed, with slow electrons at the image center, and fast electrons located towards the outer edge. Due to the relatively large C$_2$H$^-$ electron affinity (EA) of 2.969~eV~\cite{erv91}, photodetachment at 355~nm occurs close to threshold. This yields slow electrons, allowing for a low repeller voltage ($-600$~V) on the VMI lens, and high velocity resolution.

@@

The VMIs in Fig.~\ref{fig:1} were inverted using the Abel inversion methods detailed in pyAbel~\cite{hic19}, to extract the corresponding photoelectron spectra presented in Fig.~\ref{fig:1}(c). Below $27,000~$cm$^{-1}$ on the $\tilde{X} ^2\Sigma^+$ surface the C$_2$H$^-$ and C$_2$D$^-$ spectra are a close match. Both are dominated by the origin transition, shifted by $\sim10~$cm$^{-1}$ in the deuterated spectrum, with progressions involving the v$_2$($\pi$) bending and v$_3$($\sigma$) CC stretch vibrational normal modes. This includes transitions involving an odd quanta of v$_2$ bending excitation ($2^{n+1}$) which should be totally forbidden within the harmonic oscillator approximation. However, the presence of these transitions in the spectra is likely an indicator of Herzberg-Teller (HT) vibronic coupling between the ground $^2\Sigma^+$ and nearby excited $^2\Pi$ electronic surfaces, as
\begin{equation}
\Sigma^+ \otimes \pi = \Pi. 
\end{equation}

Above $27,000~$cm$^{-1}$, near the $\tilde{A}^2\Pi$ state origin, large differences are observed between the C$_2$H$^-$ and C$_2$D$^-$ photoelectron spectra. In the C$_2$H$^-$ spectrum 5 sharp peaks are observed, spaced by $\sim95~$cm$^{-1}$. However in the deuterated spectrum, one dominant peak is observed at $27,792$~cm$^{-1}$, with 3 weaker peaks centred around $27,360~$cm$^{-1}$. Unlike the structure below $27,000$~cm$^{-1}$ none of these peaks are able to be readily assigned to vibronic transitions, due to the presence of strong coupling interactions between the nearby $\Sigma^+$ and $\Pi$ surfaces.

@@

In C$_2$H this effect creates a Renner-Teller (RT) pair in the excited state, where the usually degenerate $\Pi$ surfaces separate to form two non-degenerate electronic states $\Pi^+ (2A')$ and $\Pi^-(1A'')$. This involves separating a single potential energy surface (V) into two distinct but connected surfaces (V$^+$) and (V$^-$). Due to the strong coupling along the linear axis between the electronic and vibration angular momenta of the $2A'$ and $1A''$ components of the $^2\Pi$ state, stationary states cannot be explicitly assigned to either of the $\Pi^+(2A')$ or $\Pi^-(1A'')$ electronic surfaces. Instead, they exist as a combination of both states.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{figures/Fig3}
	\caption{Adiabatic potential energy surfaces of $\tilde{X}^2\Sigma^+$, $\tilde{A}^2\Pi^-$, and $\tilde{A}^2\Pi^+$ states, calculated using the method developed by Tarroni and Carter~\cite{tar03} (E$_{1A'}$, E$_{2A'}$, and E$_{1A''}$ in Eq.~\ref{eq:teller3}).}
	\label{fig:2}
\end{figure}

The adiabatic potential energy surfaces in Figure~\ref{fig:2} were calculated using the variational method of Tarroni and Carter~\cite{car00}. This has been used to calculate a multitude of admixed vibronic levels. In the literature, over 100 possible C$_2$H vibronic states have been calculated, up to 10,000~cm$^{-1}$ above the $\tilde{X} ^2\Sigma^+$ origin. However, assigning experimental spectra has remained a challenge, partly due to the large number of calculated levels providing multiple possible assignments for each experimentally observed transition.

@@

The photoelectron spectrum of C$_2$D$^-$ was also examined using the same approach, with the calculated transitions shown in Figure~\ref{fig:3}(b) alongside the 355~nm experimental data. Again, there is excellent agreement in the intensity and calculated positions of the transitions on the $\tilde{X} ^2\Sigma^+$ state surface, including the HT coupled levels. Near the $\tilde{A} ^2\Pi$ state origin, deuteration has a large impact on the experimental photoelectron spectrum. Instead of the 5 evenly split levels observed in C$_2$H, a single dominant peak ($d$) is now observed at 3,844~cm$^{-1}$ alongside a collection of weaker peaks ($a-c$) centred around 3,400~cm$^{-1}$. Based on previous spin-orbit splitting measurements, peak d is expected to have significant $\tilde{A} ^2\Pi$ character, with peaks a-c commonly assigned to vibronic levels with predominantly HT coupled $\tilde{X} ^2\Sigma^+$ character~\cite{yan87,hsu95,chi99,wil11}. This suggests that deuteration effectively dampens the coupling interaction between the surfaces.

Like the experimental spectra, a large change is observed in the simulated spectrum upon deuteration. The calculated intensity pattern of the vibronic levels around the $\tilde{A} ^2\Pi$ origin is a good match to the experimental data, with a single intense transition observed near peak d, and 3 prominent transitions predicted around the experimental peaks $a-c$. The calculated splitting between the levels is also similar, however the relative intensity of peak $d$ appears to be slightly underestimated. %The results from Figure~\ref{fig:C2H-plot} confirm that the quasiadiabtic approach is able to accurately describe the vibronic interactions between the $\Sigma$ and $\Pi$ surfaces, including deuteration effects.

\begin{figure}[th!]
	\includegraphics[width=0.6\textwidth]{figures/266-plot.pdf}
	\caption{Photoelectron spectrum of C$_2$H$^-$ at 266~nm, showing vibrationally excited $\tilde{A}$ state structure. The simulated spectra shown in green is calculated using the quasidiabatic ansatz, and convolved with Gaussians with a FWHM of 40~cm$^{-1}$ to match the lower VMI resolution at large electron kinetic energies.}
	\label{fig:4}
\end{figure}
To investigate the effects of coupling on the higher excited vibronic levels on the$\tilde{A} ^2\Pi$ surface, the calculations were also compared to the experimental photoelectron spectrum of C$_2$H$^-$ at 266~nm, presented in Figure~\ref{fig:4}. Gaussians were fit to the transitions from Figure~\ref{fig:3}(a) with a FWHM of 40~cm$^{-1}$ to match the lower resolution of the experimental data at high electron kinetic energies. Near 3,900~cm$^{-1}$ the electronic origin of the $\tilde{A} ^2\Pi$ state is split over 5 vibronic levels, as discussed above. A similar effect is also observed for the $3^1_0$ band, which becomes split over 3 vibronic levels labelled $l$, $n$, and $o$. The calculated spectrum correctly predicts the position and intensity of peaks $n$ and $o$, but appears to strongly underestimate the intensity of peak $l$. Between the $0^0_0$ and $3^1_0$ bands a collection of weaker peaks are also observed ($f-k$). These are assigned to highly excited HT coupled $\tilde{X} ^2\Sigma^+$ transitions ($j,k$), admixed $\tilde{X}-\tilde{A}$ transitions ($h,i$) and pure $\tilde{X} ^2\Sigma^+$ transitions ($f$). Consequently, peak $f$ will possess $\Sigma$ symmetry, and should have a different anisotropy to all of the other dominant transitions above 3,800~cm$^{-1}$. The position and intensity of these highly-excited coupled transitions are well described by the simulated spectrum.

@@

Photodetachment to the ground state of C$_2$H $(\tilde{X}\,^2\Sigma^+)$ involves ejecting an electron from an $s-$like $5\sigma_g$ orbital, whereas detachment to the excited $\tilde{A}\,^2\Pi$ state occurs from a $p-$like $1\pi_u$ orbital. Therefore, the electron anisotropies may be described using the mixed $sp$ model\cite{khu14},
\begin{equation}
\beta(\epsilon)_{sp} = \frac{2Z\epsilon+2\text{A}_1\epsilon^2-4\epsilon\cos\delta_{\ell\pm 1}}{\frac{1}{\text{A}_1}+2\text{A}_1\epsilon^2+Z\epsilon}
\label{eq:beta-sanov}
\end{equation}  
where $\epsilon$ is the electron kinetic energy, $A_1$ is the $s\rightarrow p$ Hanstorp coefficient, $f$ is the percentage of $p$ orbital character,
\begin{equation}
Z = \frac{1-f}{f}\frac{8}{3}~~ \text{and} ~~ |\psi\rangle = \sqrt{1-f}|s\rangle + \sqrt{f}|p\rangle.
\end{equation}

From Eq.~(\ref{eq:beta-sanov}) it can be seen that detachment from a pure $s$ orbital will have a positive anisotropy ($\beta=+2$), whereas detachment from a pure $p$ orbital will have a negative anisotropy for electron kinetic energies $\epsilon < 2/A_1$. Therefore, measuring the anisotropy will determine the electronic character of each individual transition, which may be compared to the calculated symmetries in Figure~\ref{fig:3}. The anisotropy parameters were measured for every detachment wavelength and prominent transition in the C$_2$H$^-$ photoelectron spectra from this work, and are presented in Figure~\ref{fig:5}. Fitting Eq.~(\ref{eq:beta-sanov}) to the anisotropy parameters from $^2\Pi$ state detachment, with f=0.9 produces a Hanstorp coefficient of $A_1$=0.66(4)~eV$^{-1}$.

\begin{figure}[th!]
	\includegraphics[width=0.5\textwidth]{figures/PAD.pdf}
	\caption{Photoelectron anisotropy parameters, measured for each resolved C$_2$H$^-$ transition, across a range of detachment wavelengths. Allowed ground state transitions $\tilde{X}$ are shown in orange, Herzberg-Teller coupled transitions are shown in purple, and excited state transitions $\tilde{A}$ are shown in blue. Anisotropy curves for a mixed-$sp$ model (Eq.~\ref{eq:beta-sanov}) are shown for f=0.1 and f=0.9.}
	\label{fig:5}
\end{figure}

@@

\section{Methods}
Details of the HR-PEI spectrometer are given in Refs~\onlinecite{cav07} and~\onlinecite{dev17}. C$_2$H$^-$ ions are produced by passing a 1:1 C$_2$H$_4$:O$_2$ gas mixture through a pulse valve, which then undergoes supersonic expansion into a high-voltage discharge. Negative ions are extracted, accelerated to 500~eV, and focussed into a novel gating, bunching, and re-referencing unit~\cite{ded01}. Anions are mass separated over a 2m time-of-flight region, with the ion of interest isolated by an electrostatic gate. The ion packet is crossed with a tuneable detachment laser beam, generated from a Sunlite EX optical parametric oscillator pumped by the third harmonic of a Continuum Powerlite 9010 Nd:YAG laser. The laser produces between 10-50mJ per pulse at 10Hz, depending on whether the idler or signal beam is used. The wavelength of the laser light is measured using a HighFinesse WS7 UV wavemeter.

A velocity-map imaging lens, a modified version of the original concept of Eppink and Parker, images the detached electrons to a 75mm diameter MCP/phosphor screen detector. Events are imaged by a 2014x2048 monochrome CCD camera (PCO 2000), with each frame transferred to a computer at a 10Hz repetition rate, and processed in real time to identify individual electron events. The electron positions are centroided to a sub-pixel accuracy, then written to a data file for subsequent analysis. The velocity-map image is centred and then circularized by an angular dependent-radial scaling determined by comparing adjacent radial slice intensity profiles~\cite{gas17}. An inverse Abel transformation of the VMI, based on the algorithm of Hansen and Law~\cite{han85,pya16}, returns a slice image of the 3D electron source distribution. Absolute energy calibration of the photoelectron spectra is achieved using published measurements of species, including O$^-$~\cite{cav07} and NO$_2^-$~\cite{law19}, that have been studied under similar conditions as used for the C$_2$H$^-$ measurements.

>>>>

\documentclass[journal=jpcafh,manuscript=article,layout=onecolumn, 12pt]{achemso}
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage[T1]{fontenc}       % Use modern font encodings
\usepackage{amsmath}
\usepackage[dvipsnames]{xcolor}
% NB added command for in line cite
\newcommand{\onlinecite}[1]{\hspace{-1 ex} \nocite{#1}\citenum{#1}} 
%
\author{Benjamin~A.~Laws}
\email{b.laws@unsw.edu.au}
\affiliation{School of Chemistry, University of New South Wales, Sydney NSW 2052, Australia}
\alsoaffiliation{Research School of Physics, The Australian
	National University, Canberra ACT 2601, Australia}
\author{Zachariah~D.~Levey} 
\affiliation{School of Chemistry, University of New South Wales, Sydney NSW 2052, Australia}
\author{Andrei M. Sanov}
\affiliation{Department of Chemistry and Biochemistry, The University of Arizona, Tucson, Arizona 85721, United States}
\author{John F. Stanton}
\affiliation{Department of Chemistry, University of Florida, Gainesville, Florida 32611, United States}
\author{Timothy~W.~Schmidt} 
\affiliation{School of Chemistry, University of New South Wales, Sydney NSW 2052, Australia}
\author{Stephen~T.~Gibson}
\affiliation{Research School of Physics, The Australian
	National University, Canberra ACT 2601, Australia}
\title{Velocity Map Imaging Spectroscopy of C$_2$H$^-$ and C$_2$D$^-$: a benchmark study of vibronic coupling interactions}
\abbreviations{PES,PAD,EA,eKE,FWHM,VMI}
\begin{document} 
\begin{abstract}

@@

In this work the vibronic spectrum of the ethynyl radical C$_2$H is examined by employing High-Resolution Photoelectron Imaging (HR-PEI) in order to benchmark CFOUR~\cite{dev20}  \emph{ab-initio} vibronic-coupling calculations. C$_2$H is reported to be one of the most abundant molecules in the universe, and is the most thoroughly studied of the C$_{2n}$H species~\cite{wil91,hei99,for10}. Many different experimental techniques have been employed to try and resolve the complex vibronic spectrum, including electron spin resonance~\cite{coc64,jin85}, laser magnetic resonance~\cite{bro88,pfe96,sch98}, microwave and milli-meter wave spectroscopy~\cite{sas81,got83,end89,mul00}, infrared (matrix isolation~\cite{she87,jac87,for95} and Fourier Transform emission~\cite{ver88}) spectroscopy, laser induced fluorescence spectroscopy~\cite{hsu93,hsu95,chi99} and photoelectron spectroscopy~\cite{erv91,tay98,zho07}. C$_2$H has also received extensive theoretical attention to try and understand the experimental results, however the large number of possible vibronic levels means this has proven to be a challenge~\cite{per90,per91,per91b,per91c,per92,car00,tar03,tar04,for10,sta21}. In this work we demonstrate how construction of a quasidiabatic Hamiltonian allows for the strength of vibronic interactions between coupled surfaces near a conical intersection to be calculated, in order to simulate electronic and vibronic spectra. By employing anion HR-PEI both the $^2\Sigma^+$ and $^2\Pi$ surfaces are mapped out on an equal footing from the anion $\tilde{X}^1\Sigma^+$ state, allowing a direct comparison with \emph{ab-initio} modelling.

\section{Results and Analysis}
Ethynyl ions were produced in a pulsed-jet discharge of pure C$_2$H$_4$ gas, and subsequently mass isolated via time of flight. Electrons were detached using a tuneable Continuum Sunlite Ex Optical Parametric Oscillator (OPO) pumped with the third harmonic of a Nd:YAG laser. High electron counts could also be obtained by using the third and fourth harmonics of the Nd:YAG laser directly. Detached electrons are mapped onto a micro channel plate detector using a Velocity Map Imaging (VMI) lens. An illustrative VMI of $\sim$4 million electrons collected from 355~nm (3.49~eV) photodetachment of C$_2$H$^-$ is shown in Fig.~\ref{fig:1}(a). The electrons are distributed radially according to their speed, with slow electrons at the image center, and fast electrons located towards the outer edge. Due to the relatively large C$_2$H$^-$ electron affinity (EA) of 2.969~eV~\cite{erv91}, photodetachment at 355~nm occurs close to threshold. This yields slow electrons, allowing for a low repeller voltage ($-600$~V) on the VMI lens, and high velocity resolution.

@@

The VMIs in Fig.~\ref{fig:1} were inverted using the Abel inversion methods detailed in PyAbel~\cite{hic19}, to extract the corresponding photoelectron spectra presented in Fig.~\ref{fig:1}(c). Below $27,000~$cm$^{-1}$ on the $\tilde{X} ^2\Sigma^+$ surface the C$_2$H$^-$ and C$_2$D$^-$ spectra are a close match. Both are dominated by the origin transition, shifted by $\sim10~$cm$^{-1}$ in the deuterated spectrum, with progressions involving the v$_2$($\pi$) bending and v$_3$($\sigma$) CC stretch vibrational normal modes. This includes transitions involving an odd quanta of v$_2$ bending excitation ($2^{n+1}$) which should be totally forbidden within the harmonic oscillator approximation. However, the presence of these transitions in the spectra is likely an indicator of Herzberg-Teller (HT) vibronic coupling between the ground $^2\Sigma^+$ and nearby excited $^2\Pi$ electronic surfaces, as
\begin{equation}
\Sigma^+ \otimes \pi = \Pi. 
\end{equation}

Above $27,000~$cm$^{-1}$, near the $\tilde{A}^2\Pi$ state origin, significant differences are observed between the C$_2$H$^-$ and C$_2$D$^-$ photoelectron spectra. In the C$_2$H$^-$ spectrum 5 sharp peaks are observed, spaced by $\sim95~$cm$^{-1}$. However in the deuterated spectrum, one dominant peak is observed at $27,792$~cm$^{-1}$, with 3 weaker peaks centred around $27,360~$cm$^{-1}$. Unlike the structure below $27,000$~cm$^{-1}$ none of these peaks are able to be readily assigned to vibronic transitions, due to the presence of strong coupling interactions between the nearby $\Sigma^+$ and $\Pi$ surfaces.

@@

In C$_2$H this effect creates a Renner-Teller (RT) pair in the excited state, where the usually degenerate $\Pi$ surfaces separate to form two non-degenerate electronic states $\Pi^+ (2A')$ and $\Pi^-(1A'')$. This involves separating a single potential energy surface (V) into two distinct but coupled surfaces (V$^+$) and (V$^-$). Due to the strong coupling along the linear axis between the electronic and vibration angular momenta of the $2A'$ and $1A''$ components of the $^2\Pi$ state, stationary states cannot be explicitly assigned to either of the $\Pi^+(2A')$ or $\Pi^-(1A'')$ electronic surfaces. Instead, they exist as a combination of both states.

@@

\begin{figure}
	\centering
	\includegraphics[width=0.5\textwidth]{figures/Fig3}
	\caption{Adiabatic potential energy surfaces of $\tilde{X}^2\Sigma^+$, $\tilde{A}^2\Pi^-$, and $\tilde{A}^2\Pi^+$ states, calculated using the parameters calculated by Tarroni and Carter~\cite{tar03} (E$_{1A'}$, E$_{2A'}$, and E$_{1A''}$ in Eq.~\ref{eq:teller3}).}
	\label{fig:2}
\end{figure}

The adiabatic potential energy surfaces in Figure~\ref{fig:2} were evaluated from the variational parameters of Tarroni and Carter~\cite{car00}. Their technique calculates a multitude of admixed vibronic levels. In the literature, over 100 possible C$_2$H vibronic states have been calculated, up to 10,000~cm$^{-1}$ above the $\tilde{X} ^2\Sigma^+$ origin. However, assigning experimental spectra has remained a challenge, partly due to the large number of calculated levels providing multiple possible assignments for each experimentally observed transition.

@@

The photoelectron spectrum of C$_2$D$^-$ was also examined using the same approach, with the calculated transitions shown in Figure~\ref{fig:3}(b) alongside the 355~nm experimental data. Again, there is excellent agreement in the intensity and calculated positions of the transitions on the $\tilde{X} ^2\Sigma^+$ state surface, including the HT coupled levels. Near the $\tilde{A} ^2\Pi$ state origin, deuteration has a large impact on the experimental photoelectron spectrum. Instead of the 5 evenly split levels observed in C$_2$H, a single dominant peak ($d$) is now observed at 3,844~cm$^{-1}$ alongside a collection of weaker peaks ($a-c$) centred around 3,400~cm$^{-1}$. Based on previous spin-orbit splitting measurements, peak d is expected to have significant $\tilde{A} ^2\Pi$ character, with peaks a-c commonly assigned to vibronic levels with predominantly HT coupled $\tilde{X} ^2\Sigma^+$ character~\cite{yan87,hsu95,chi99,wil11}. This suggests that deuteration effectively dampens the coupling interaction between the surfaces, as has been observed for HCCH~\cite{dev17}.

In agreement with the experimental spectra, a large change is observed in the simulated spectrum upon deuteration. The calculated intensity pattern of the vibronic levels around the $\tilde{A} ^2\Pi$ origin is also a good match to the experimental data, with a single intense transition observed near peak d, and 3 prominent transitions predicted around the experimental peaks $a-c$. The calculated splitting between the levels is also similar, however the relative intensity of peak $d$ appears to be slightly underestimated. %The results from Figure~\ref{fig:C2H-plot} confirm that the quasiadiabtic approach is able to accurately describe the vibronic interactions between the $\Sigma$ and $\Pi$ surfaces, including deuteration effects.

\begin{figure}[th!]
	\includegraphics[width=0.6\textwidth]{figures/266-plot.pdf}
	\caption{Photoelectron spectrum of C$_2$H$^-$ at 266~nm, showing vibrationally excited $\tilde{A}$ state structure. The simulated spectra shown in green is calculated using the quasidiabatic ansatz, and convolved with Gaussians with a FWHM of 40~cm$^{-1}$ to match the lower VMI resolution at large electron kinetic energies.}
	\label{fig:4}
\end{figure}
To investigate the effects of coupling on the higher excited vibronic levels on the $\tilde{A} ^2\Pi$ surface, the calculations were also compared to the experimental photoelectron spectrum of C$_2$H$^-$ at 266~nm, presented in Figure~\ref{fig:4}. Gaussians were fit to the transitions from Figure~\ref{fig:3}(a) with a FWHM of 40~cm$^{-1}$ to match the lower resolution of the experimental data at high electron kinetic energies. Near 3,900~cm$^{-1}$ the electronic origin of the $\tilde{A} ^2\Pi$ state is split over 5 vibronic levels, as discussed above. A similar effect is also observed for the $3^1_0$ band, which becomes split over 3 vibronic levels labelled $l$, $n$, and $o$. The calculated spectrum correctly predicts the position and intensity of peaks $n$ and $o$, but appears to strongly underestimate the intensity of peak $l$. Between the $0^0_0$ and $3^1_0$ bands a collection of weaker peaks are also observed ($f-k$). These are assigned to highly excited HT coupled $\tilde{X} ^2\Sigma^+$ transitions ($j,k$), admixed $\tilde{X}-\tilde{A}$ transitions ($h,i$) and pure $\tilde{X} ^2\Sigma^+$ transitions ($f$). Consequently, peak $f$ will possess $\Sigma$ symmetry, and should have a different anisotropy to all of the other dominant transitions above 3,800~cm$^{-1}$. The position and intensity of these highly-excited coupled transitions are well described by the simulated spectrum.

@@

Photodetachment to the ground state of C$_2$H $(\tilde{X}\,^2\Sigma^+)$ involves ejecting an electron from an $s-$like $\sigma$ orbital (approximately $5\sigma_g$ in symmetry character), whereas detachment to the excited $\tilde{A}\,^2\Pi$ state occurs from a $p-$like $\pi$ orbital (approximately $1\pi_u$ in character). Therefore, the electron anisotropies may be described using the mixed $s-p$ model\cite{khu14},
\begin{equation}
\beta_{sp}(\epsilon) = \frac{2(1-\gamma_p)B_1\epsilon + \gamma_p(2A_1^2\epsilon^2-4A_1\epsilon\cos\delta_{2,0})}
{(1-\gamma_p)B_1\epsilon+\gamma_p(1+2A_1^2\epsilon^2)}
\label{eq:beta-sp}
\end{equation}
where $\epsilon$ is the electron kinetic energy and $\gamma_p$ is the fraction of $p$ character of the detachment orbital described as
\begin{equation}
|\psi\rangle = \sqrt{1-\gamma_p}|s\rangle + \sqrt{\gamma_p}|p\rangle.
\end{equation}
$A_1$ and $B_1$ in Eq.~(\ref{eq:beta-sp}) are the generalised Hanstorp~\cite{han89} coefficients describing the assumed Wigner-like~\cite{wig48} relative scalings of the radial transition dipole matrix elements for different allowed detachment channels. Specifically, $A_1\epsilon$ describes the energy-dependent ratio of the $p\rightarrow d$ and $p\rightarrow s$ transition amplitudes, while $B_1\epsilon$ corresponds to the $s\rightarrow p$ and $p\rightarrow s$ cross-section ratio~\cite{khu14}. It can be shown that under certain approximations $B_1/A_1 = 8/3$~\cite{san13}. Finally, $\delta_{2,0}$ in Eq.~(\ref{eq:beta-sp}) is the phase shift between the $s$ and $d$ partial waves, which in most cases of anion photodetachment is assumed to be small, corresponding to $\cos\delta_{2,0}\approx1$.

From Eq.~(\ref{eq:beta-sp}) it can be seen that detachment from a pure $s$ orbital ($\gamma_p=0$) will have a positive anisotropy ($\beta = +2$), whereas detachment from a pure $p$ orbital ($\gamma_p=1$) will have a negative anisotropy for electron kinetic energies $\epsilon < 2/A_1$. Therefore, measuring the anisotropy can help determine the electronic character of each individual transition, which may be compared to the calculated symmetries in Figure~\ref{fig:3}. The anisotropy parameters were measured at every detachment wavelength for prominent transitions in the C$_2$H$^-$ photoelectron spectra from this work, and are presented in Figure~\ref{fig:5}. Fitting Eq.~(\ref{eq:beta-sp}) to the anisotropy parameters from $^2\Pi$ state detachment, with $\gamma_p = 0.9$ produces a Hanstorp coefficient of $A_1=0.66(4)~$eV$^{-1}$.

%\begin{equation}
%\beta(\epsilon)_{sp} = \frac{2Z\epsilon+2\text{A}_1\epsilon^2-4\epsilon\cos\delta_{\ell\pm 1}}{\frac{1}{\text{A}_1}+2\text{A}_1\epsilon^2+Z\epsilon}
%\label{eq:beta-sanov}
%\end{equation}  
%where $\epsilon$ is the electron kinetic energy, $A_1$ is the $s\rightarrow p$ Hanstorp coefficient, $f$ is the percentage of $p$ orbital character,
%\begin{equation}
%Z = \frac{1-f}{f}\frac{8}{3}~~ \text{and} ~~ |\psi\rangle = \sqrt{1-f}|s\rangle + \sqrt{f}|p\rangle.
%\end{equation}

%From Eq.~(\ref{eq:beta-sanov}) it can be seen that detachment from a pure $s$ orbital will have a positive anisotropy ($\beta=+2$), whereas detachment from a pure $p$ orbital will have a negative anisotropy for electron kinetic energies $\epsilon < 2/A_1$. Therefore, measuring the anisotropy will determine the electronic character of each individual transition, which may be compared to the calculated symmetries in Figure~\ref{fig:3}. The anisotropy parameters were measured for every detachment wavelength and prominent transition in the C$_2$H$^-$ photoelectron spectra from this work, and are presented in Figure~\ref{fig:5}. Fitting Eq.~(\ref{eq:beta-sanov}) to the anisotropy parameters from $^2\Pi$ state detachment, with f=0.9 produces a Hanstorp coefficient of $A_1$=0.66(4)~eV$^{-1}$.

\begin{figure}[th!]
	\includegraphics[width=0.5\textwidth]{figures/PAD.pdf}
	\caption{Photoelectron anisotropy parameters, measured for each resolved C$_2$H$^-$ transition, across a range of detachment wavelengths. Allowed ground state transitions $\tilde{X}^2\Sigma^+$ are shown in orange, Herzberg-Teller coupled transitions are shown in purple, and excited state transitions $\tilde{A} ^2\Pi$ are shown in blue. Anisotropy curves for a mixed-$sp$ model (Eq.~\ref{eq:beta-sp}) are shown for $\gamma_p=0.1$ and $\gamma_p=0.9$.}
	\label{fig:5}
\end{figure}

@@

\section{Methods}
Details of the HR-PEI spectrometer are given in Refs~\onlinecite{cav07} and~\onlinecite{dev17}. C$_2$H$^-$ ions are produced by passing a 1:1 C$_2$H$_4$:O$_2$ gas mixture through a pulse valve, which then undergoes supersonic expansion into a high-voltage discharge. Negative ions are extracted, accelerated to 500~eV, and focussed into a novel gating, bunching, and re-referencing unit~\cite{ded01}. Anions are mass separated over a 2m time-of-flight region, with the ion of interest isolated by an electrostatic gate. The ion packet is crossed with a tuneable detachment laser beam, generated from a Continuum Sunlite EX optical parametric oscillator pumped by the third harmonic of a Continuum Powerlite 9010 Nd:YAG laser. The laser produces between 10-50mJ per pulse at 10Hz, depending on whether the idler or signal beam is used. The wavelength of the laser light is measured using a HighFinesse WS7 UV wavemeter.

A velocity-map imaging lens (modified version of the original concept of Eppink and Parker) images the detached electrons to a 75mm diameter MCP/phosphor screen detector. Events are imaged by a 2048x2048 monochrome CCD camera (PCO 2000), with each frame transferred to a computer at a 10Hz repetition rate, and processed in real time to identify individual electron events. The electron positions are centroided to a sub-pixel accuracy, then written to a data file for subsequent analysis. The velocity-map image is centred and then circularized by an angular dependent-radial scaling determined by comparing adjacent radial slice intensity profiles~\cite{gas17}. An inverse Abel transformation of the VMI, based on the algorithm of Hansen and Law~\cite{han85,hic19}, returns a slice image of the 3D electron source distribution. Absolute energy calibration of the photoelectron spectra is achieved using published measurements of species, including O$^-$~\cite{cav07} and NO$_2^-$~\cite{law19}, that have been studied under similar conditions as used for the C$_2$H$^-$ measurements.
